Adaptive Waveform Inversion for Transmitted Wavefields

### 2024-10-05 repro/papers/awi/paper.tex
more

@@

Each acoustic source is modeled an isotropic point radiator
with known location $\bx_s\in \bR^3$. These
sources generate acoustic waves traveling through Euclidean space. The
data are samples of these waves at receivers, modeled as isotropic
point sensors, recording pressure at points
${\bf x}_r \in \bR^3$. The recording takes place over a time interval
$[0,t_d]$ (also idealized as independent of $\bx_s$ and
$\bx_r$) . The collection $X_{sr}$ of source-receiver pairs
$(\bx_s,\bx_r)$ for which data is recorded should not intersect the
diagonal - that is, no receiver should be co-located with a source for
which it is active. The set of source locations $X_s = \{\bx_s: (\bx_s,\bx_r) \in X_{sr}
\mbox{ for some } \bx_r \in \bR^3\}$ will also be useful.

In practice, recording takes place over a finite time interval
$[0,t_d]$. For convenience, assume that this interval is the same for
all receivers. Let $\phi \in C_0^{\infty}(0,t_d)$ be a window function
supported in this interval. Additional constraints on $t_d$ and $\phi$
will be be specified below.

The pressure and velocity fields $p({\bf x},t;{\bf x}_s)$,
${\bf v}({\bf x},t;{\bf x}_s)$ for the source location ${\bf x}_s$
depend on the bulk modulus $\kappa({\bf x})$, buoyancy $\beta({\bf
  x})$ (reciprocal of the density $\rho({\bf x})$), and source wave
form $f(t)$ through the acoustic system
\begin{eqnarray}
  \label{eqn:awe}
 \frac{\partial p}{\partial t} & = &- \kappa \nabla \cdot {\bf v} +
                                    f(t) \delta({\bf x}-{\bf x}_s); \nonumber \\
\frac{\partial {\bf v}}{\partial t} & = & - \beta \nabla p; \\ 
p, {\bf v} & = & 0 \mbox{ for }  t \ll 0.
\end{eqnarray}
For simplicity, I assume that the system \ref{eqn:awe} holds over all of Euclidean
space-time $(\bx,t) \in \bR^3 \times \bR$, and ignore physical
boundaries such as the Earth's surface.

As mentioned in the introduction, I will assume that the coefficients
$\rho$ and $\kappa$ are bounded and smooth:
\begin{equation}
  \label{eqn:coeff}
  \log \rho, \log \kappa \in C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
\end{equation}
Then the symmetric hyperbolic system \ref{eqn:awe} has a distribution
solution if $f \in \cal{E}'(\bR)$ (for example \cite{Lax:PDENotes}). The Propagation of
Singularities Theorem (for example, \cite{Tay:81}, Theorem VI.2.1)
implies that if $f \in \cal{E}'(\bR)$ and $(\bx_s,\bx_r) \in X_{sr}$,
then the trace $p(\bx_r, \cdot;\bx_s)$ on $\{(\bx_r,t): t \in \bR\}$ is well-defined,
as a distribution vanishing for sufficiently large
negative $t$. Define the {\em modeling operator}
\begin{equation}
  \label{eqn:modop}
  F: (C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3))^2 
  \rightarrow X_{sr} \times {\cal D}'(\bR)
\end{equation}
by
\begin{equation}
  \label{eqn:modopdef}
  F[\kappa,\rho](\bx_r,\bx_s) = \phi p(\bx_r,\cdot;\bx_s).
\end{equation}
For sake of brevity, define the model vector $m=(\kappa,\rho)$.

For the purposes of this paper, the source wavelet $f$ will be assumed
known, and to satisfy
\begin{itemize}
\item[F1. ] $f \in H^1_0(\bR)$,
\item[F2. ] $f$ is odd, $f(t) = -f(-t)$.
\end{itemize}
In fact, even if actual acoustic sources are well-approximated by
isotropic point radiators, their wave forms are only roughly known
{\em a priori} and may reasonably be included as parameters to be
estimated by inversion of the data. On the other hand, in applications
to seismology, medical imaging, and other wave imaging technologies,
the source wave form $f$ is actually the product of signal processing
aimed to shape the data $F[m]$ into an approximation of data that
would have been predicted using a prescribed $f$. That is, the wavelet
is to some extent {\em chosen}. This circumstance justifies the
constraints F1, F2 placed on $f$.

The density $\rho$ will be presumed spatially homogeneous, or at least
to have the same value at all of the source positions $\bx_s\in X_s$:
introduce the {\em source density} $\rho_s > 0$ for which
\begin{equation}
  \label{eqn:rho}
  \rho(\bx_s) = \rho_s\mbox{ for any } \bx_s \in X_s.
\end{equation}
In some settings, eg. the earth's
upper crust, density exhibits substantially less variance than the
bulk modulus, Also, for some acquisition designs, such as marine towed
streamers and some medical ultrasonic technologies, sources are
immersed in water and thus in positions with approximately the same
densities.

In any event, for the present discussion, $f$ is regarded
as known and subject to F1 and F2, and $\rho$ is subject to the constraint \ref{eqn:rho}.

In this setting, acoustic FWI means: given the set of
source-receiver position pairs $X_{sr}$, source wavelet
$f$,  source density $\rho_s$, and data traces
$d:  X_{sr} \rightarrow \cal{D}(\bR)$, find $\kappa$ and $\rho$
(subject to the constraint \ref{eqn:rho}) so that the model
$m=(\kappa,\rho)$ satisfies $F[m] \approx d$ in some sense. If
$d$ and $F[m]$ lie in a normed subspace of $\cal{D}(\bR)$ for each
$(\bx_s,\bx_r) \in X_{sr}$, then the
norm provides a straightforward candidate for this sense of
approximation.

\subsection{Hadamard's Decomposition}
The Hadamard decomposition of the acoustic Green's function provides a
basis for a norm-based formulation of FWI \cite[]{Friedlander:75,Qian:JCP24}. As presented by
\cite{Friedlander:75}, for instance, this construction applies to
second-order wave equations, but the system \ref{eqn:awe} is
equivalent fo a second-order problem: differentiate the first equation
with respect to $t$, eliminate $\bv$ via the second equation, and
rearrange to obtain
\begin{equation}
  \label{eqn:awe2}
  \frac{\rho}{\kappa}\frac{\partial^2 p}{\partial t^2} - \nabla^2 p -
  \nabla \log \rho \cdot \nabla p = \rho \frac{\partial f}{\partial t}
  \delta(\bx-\bx_s)
\end{equation}
The pair $(p,\bv)$ is a distribution solutions of the system
\ref{eqn:awe} if and only if $p$ is a solution of equation
\ref{eqn:awe2} and vis-versa (with $\bv$ constructed from $p$ in the
obvious way, preserving causality).

Hadamard's construction is local, in both space-time and in the set of
acoustic models. A {\em normal neighborhood} of
$\bx_s \in \bR^3$ is an open subset of $\bR^3$ 
within which every point is connected to $\bx_s$ by  a unique ray of
geometric acoustics (a geodesic for the Riemannian metric associated
to the equation \ref{eqn:awe2}) (\cite{Friedlander:75}, p. 16). In the
following, 
$M \subset (C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3))^2$ and $\Omega
\subset \bR^3$ are bounded open sets satisfying 
\begin{itemize}
  \item[G1. ] for every $(\bx_s,\bx_r) \in X_{sr}$, $\bx_s, \bx_r \in \Omega$;
  \item[G2. ] for every source point $\bx_s\in X_{s}$ and $(\kappa,\rho) \in M$, $\Omega$ is a
    normal neighborhood of $\bx_s$.
  \item[G3. ] for every source point $\bx_s \in X_{s};$
    $(\kappa,\rho) \in M$, ray ${\bf X}: [0,t_d] \rightarrow \bR^3$
    with ${\bf X}(0)=\bx_s$, and
    $t \in [0,t_d]$, either $\{{\bf X}(s): 0\le s \le t\} \subset \Omega$
    or $\{{\bf X}(s): t < s < t_d\} \subset \bR^3 \setminus \Omega$.
  \end{itemize}
The last condition means that once a ray leaves $\Omega$ within the
recording time interval $[0,t_d]$, it does not return to $\Omega$
for $t<t_d$.

Note that for any $X_{sr}$ there exist
non-empty open sets $M$ and $\Omega$ satisfying these conditions: any convex open set
$\Omega$ satisfying condition G1 satisfies conditions G2 and G3 for constant
($\bx$-independent) $\kappa$ and $\rho$, and conditions G2 and G3 are open in
the sense of the $C^{\infty}$ and $L^{\infty}$ topologies.

Under the conditions just described, the construction described by
\cite{Friedlander:75} produces for each $m \in M$ functions $\tau[m]$ (travel time) and $a[m]$
(geometric amplitude) on $X_s \times \Omega$ and $b[m]$ (remainder kernel) on $X_s \times \Omega
\times \bR$, satisfying
\begin{itemize}
\item[H1. ] $\tau[m], a[m],$ and $b[m]$ are smooth off the diagonal,
that is, $\tau[m](\bx_s,\cdot) \in C^{\infty}(\Omega \setminus \{\bx_s\})$
and similarly for $a[m],b[m]$;
\item[H2. ] there exists $A>0$ so that for any $(\kappa,\rho) \in M$
  and $(\bx_s,\bx_r) \in X_{sr}$,
  $|\log a[m](\bx_s,\bx_r)| < A$;
\item[H3. ] for any $t>0$, there is $B\ge 0$ so that for
$(\bx_s,\bx_r) \in X_{sr}, -\infty < s \le t$, and $(\kappa,\rho) \in
M$, $|b[m](\bx_s,\bx_r,s)| +
|\partial_tb[m](\bx_s,\bx_r,s)| \le B$.
\end{itemize}

The Hadamard
decomposition of the Green's function $G$, i.e. the solution of the
equation \ref{eqn:awe2} for $f = H/\rho_s$, is (suppressing the
dependence on $m \in M$ from the notation)::
\begin{equation}
  \label{eqn:green}
  G(\bx_s,\bx,t) = a(\bx_s,\bx)\delta(t - \tau(\bx_s,\bx)) +
  b(\bx_s,\bx,t)H(t - \tau(\bx_s,\bx))
\end{equation}
Under conditions G1 - G3 above, this decomposition is the restriction of
the Greeen's function to $\Omega \times (-\infty,t_d]$. Accordingly,
\[
   F[m](\bx_s,\bx_r)(t) =
   \phi(t)a(\bx_s,\bx_r)\rho_s\frac{df}{dt}(t-\tau(\bx_s,\bx_r))
 \]
\begin{equation}
  \label{eqn:fwdgreen}
 + \phi(t)\rho_s\int
  \,ds \, \frac{df}{dt}(t-s) b(\bx_s,\bx_r,s)H(s-\tau(\bx_s,\bx_r))
\end{equation}

\subsection{FWI via Least Squares}
It follows from the Hadamard decomposition 
(equation \ref{eqn:fwdgreen}) that $F[m]$ is
square-integrable. Indeed, for each $(\bx_s,\bx_r) \in X_{sr}$, the first term on the RHS of equation \ref{eqn:fwdgreen} is clearly
square-integrable, since $f \in H^1_0(\bR)$.
An $L^2$ bound on the second (remainder) term follows from the
pointwise bounds on $b$ described in the last section and Young's
inequality. In fact, there exists $C>0$ so that for $(\kappa,\rho) \in
M$,
\begin{equation}
  \label{eqn:fwdnorm}
  \|F[m]\|_d \le C
  \left\|\frac{df}{dt}\right\|_{L^2(\bR)}.
\end{equation}
in which the left-hand side uses the abbreviation
\[
  \| \cdot \|_d = \| \cdot \|_{L^2(X_{sr} \times \bR)}
\]

This observation justfies posing FWI as a least-squares problem: given $d \in
L^2(X_{sr}\times \bR)$, $f \in H^1_0(\bR)$, and $\rho_s>0$, find $m = (\kappa,\rho)$ to minimize
\begin{equation}
  \label{eqn:fwi}
  J_{\rm FWI}[m;d]= \frac{1}{2}\|F[m]-d\|_d^2.
\end{equation}
Practical versions of least-squares FWI include various types of regularization and
imposition of constraints on $m$, see the references cited in the
Introduction.
%As mentioned there, local iterative optimization methods
%are the only practical approach to minimizing $J_{\rm FWI}$ due to the
%computational cost of approximating the forward map $F$ and related
%operators, but these tend to stagnate at highly sub-optimal
%(``cycle-skipped'') estimates of $m$ unless the initial estimate is of
%quite high quality, often available only at considerable cost in data
%collection and processing.

@@

Note that in the representation \ref{eqn:fwdgreen}, only the
derivative $df/dt$ appears. For convenience, introduce
\begin{equation}
  \label{eqn:wavelet}
  w = \rho_s \frac{df}{dt}
\end{equation}
From the assumed
properties of $f$, $w \in L^2(\bR)$ and has compact support, has zero
mean, and is
even in $t$. Since $f$ with the properties F1 - F3 corresponds
biuniquely with $w$ with the properties just mentioned, I will also
refer to $w$ as ``the wavelet'', and view $F$ as parametrized by $w$.

Introduce asymptotics in this setting by assuming that $w = \wl$ is chosen
from a family $\{\wl:\}$ indexed by the dimensionless parameter
$0 <\lambda \le 1\}$, and defined by
\begin{equation}
  \label{eqn:wfam}
  \wl(t) = \lambda^{-1/2}w_1(t/\lambda).
\end{equation}
The  ``mother wavelet'' $w_1$ is square-integrable of compact support
even, and mean-zero. Define $t_w$ to be the diameter of the support:
\begin{equation}
  \label{eqn:wdiam}
  \mbox{supp }w_1 = [-t_w,t_w].
\end{equation}
All wavelets $\wl$ in this family
inherit these properties from $w_1$, along with its dimensions
(energy/time$^2$). It follows from these assumptions that the Fourier transform $\fwl$ is
real-valued, even, of class $H^1(\bR)$. The corresponding source waveform $f_{\lambda}$ is
given by
\begin{equation}
  \label{eqn:flambda}
  f_{\lambda}(t) = \int_{-\infty}^t\wl =
  \lambda^{1/2}f_1\left(\frac{t}{\lambda}\right).
\end{equation}
Note that
\begin{equation}
  \label{eqn:lambdanorms}
  \|\wl\|_{L^2(\bR)} = \|w_1\|_{L^2(\bR)},\,\|f_{\lambda}\|_{L^2(\bR)}
  = \lambda \|f_1\|_{L^2(\bR)}.
\end{equation}

@@

The modeling operator with wavelet $\wl$ is denoted $\Fl$. From the
decomposition \ref{eqn:fwdgreen}, $\Fl$ is the sum of two terms, the second of which is
\[
   \phi(t)\rho_s\int
   \,ds \, \frac{df_{\lambda}}{dt}(t-s) b(\bx_s,\bx_r,s)H(s-\tau(\bx_s,\bx_r))
\]
\[
  =\phi(t)\rho_s
  f_{\lambda}(t-\tau(\bx_s,\bx_r))b(\bx_s,\bx_r,\tau(\bx_s,\bx_r))
\]
\[
    -
    \phi(t)\rho_s\int
   \,ds \, f_{\lambda}(t-s) (\partial_t
   b(\bx_s,\bx_r,s))H(s-\tau(\bx_s,\bx_r))
\]
The $L^2$ norms of the summands are bounded by a multiple of
$\lambda$, uniform over $(\bx_s,\bx_r) \in X_{sr}, m \in M$, thanks to
the norm scaling \ref{eqn:lambdanorms}, Young's inequality, and conclusion H3.

I will write ``$+ \Ol$'' to mean an  error uniformly bounded in $L^2$ by
a multiple of $\lambda$ over $m \in M$. Thus
\begin{equation}
  \label{eqn:fwdapprox}
  \Fl[m](\bx_s,\bx_r) = \phi a(\bx_s,\bx_r)\wl(t-\tau(\bx_s,\bx_r)) +
  \Ol
\end{equation}
\subsection{AWI via Optimization}

AWI is based an {\em extension} of the modeling operator $F$, that is,
an operator ${\bar F}$ with larger domain, coinciding with $F$ on its domain..
The extended modeling operator ${\bar F}$ maps extended models
$\bar{m}=(\kappa,\rho,\bar{w})$ incorporating extended sources
$\bar{w}({\bf x}_r,t;{\bf x}_s)$ to the same sampling of the pressure
field. That is, the extended source depends on the receiver location
as well as the source location - so there is one acoustic system for
each source {\em and} receiver position - a lot of wave equations! If all
of the wavelets for each source are the same, that is, ${\bar w}({\bf
  x}_r,t;{\bf x}_s) = w(t;{\bf x}_s)$ is independent of receiver
position, then ${\bar F}[\bar{m}] = F[m]$. That is, $F$ is a special
case, or restriction, of ${\bar F}$, so ${\bar F}$ is an extension of
$F$. This is the {\em source-receiver} extension, in the terminology of \cite{HuangSymes2015SEG}.

>>>>

The conclusions of this study depend on a number of limitations on the
acoustic model of wave propagation and the formulation of the related
inverse problem. The principal constraints are:
\begin{itemize}
  \item energy sources are modeled as isotropic point radiators, with
    known time dependence vanishing outside a finite time window;
  \item mechanical parameter fields (coefficients in the acoustic wave
    equation) are smooth, that is, infinitely differentiable as
    functions of position, and the material density is known and
    position-independent: only the bulk modulus (or wave velocity) is
    to be determined;
  \item waves propagate throughout Euclidean space-time - that is, no
    physical boundaries are modeled;
  \item the data of the inverse problem are the restrictions of the
    pressure field to a finite number of points in space (receiver
    positions), for a finite number of source positions and all time;
  \item local energy decays exponentially - that is, the pressure and
    all of its derivatives tend to zero at an exponential rate as $t
    \rightarrow \infty$,
    uniformly in any bounded spatial domain.
  \end{itemize}
  These constraints are discussed in detail in the next section. Some
  could certainly be relaxed. For instance, exponential decay of local
  energy implies that the finite time duration of actual measured data
  can be treated as a perturbation of the infinite time data setting
  posed here. However these hypotheses enable a simple and clear
  analysis of AWI, and seem a reasonable starting point for further
  investigation.

@@

Each acoustic source is modeled an isotropic point radiator
with known location $\bx_s\in \bR^3$. These
sources generate acoustic waves traveling through Euclidean space. The
data are samples of these waves at receivers, modeled as isotropic
point sensors, recording pressure at points
${\bf x}_r \in \bR^3$. The collection $X_{sr}$ of source-receiver pairs
$(\bx_s,\bx_r)$ for which data is recorded should not intersect the
diagonal - that is, no receiver should be co-located with a source for
which it is active. The set of source locations $X_s = \{\bx_s: (\bx_s,\bx_r) \in X_{sr}
\mbox{ for some } \bx_r \in \bR^3\}$ will also be useful.

The pressure and velocity fields $p({\bf x},t;{\bf x}_s)$,
${\bf v}({\bf x},t;{\bf x}_s)$ for the source location ${\bf x}_s$
depend on the bulk modulus $\kappa$, buoyancy $\beta$
(reciprocal of the density $\rho$), and source wave
form $f(t)$ through the acoustic system
\begin{eqnarray}
  \label{eqn:awe}
 \frac{\partial p}{\partial t} & = &- \kappa \nabla \cdot {\bf v} +
                                    f(t) \delta({\bf x}-{\bf x}_s); \nonumber \\
\frac{\partial {\bf v}}{\partial t} & = & - \beta \nabla p; \\ 
p, {\bf v} & = & 0 \mbox{ for }  t \ll 0.
\end{eqnarray}
For simplicity, I assume that the system \ref{eqn:awe} holds over all of Euclidean
space-time $(\bx,t) \in \bR^3 \times \bR$, and ignore physical
boundaries such as the Earth's surface. Likewise for simplicity, I
assume in this study that the material density $\rho$ is spatially
homogeneous and known. In some settings, eg. the earth's
upper crust, the material density $\rho$ exhibits substantially less variance than the
bulk modulus, Density is however spatially variable in most geological
and biomedical materials, and could be included amongst the parameters
to be estimated.

The parameter to be determined in the inverse problem studied here is
the bulk modulus $\kappa$. It is presumed is bounded and smooth:
\begin{equation}
  \label{eqn:coeff}
  \log \kappa \in C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
\end{equation}
Under these hypotheses, the symmetric hyperbolic system \ref{eqn:awe}
has a unique distribution
solution if $f \in \cal{E}'(\bR)$ (for example \cite{Lax:PDENotes},
Theorem 7.4). The Propagation of
Singularities Theorem (for example, \cite{Tay:81}, Theorem VI.2.1)
implies that if $f \in \cal{E}'(\bR)$ and $(\bx_s,\bx_r) \in X_{sr}$,
then the trace $p(\bx_r, \cdot;\bx_s)$ on $\{(\bx_r,t): t \in \bR\}$ is well-defined,
as a distribution vanishing for sufficiently large
negative $t$. Define the {\em modeling operator}
\begin{equation}
  \label{eqn:modop}
  F: C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
  \rightarrow X_{sr} \times {\cal D}'(\bR)
\end{equation}
by
\begin{equation}
  \label{eqn:modopdef}
  F[\kappa]= p|_{X_{sr} \times \bR}
\end{equation}

In this setting, acoustic FWI means: given the set of
source-receiver position pairs $X_{sr}$, source wavelet
$f$ subject to constraints F1 and F2,  density $\rho$, and data traces
$d:  X_{sr} \rightarrow \cal{D}(\bR)$, find $\kappa$ (or equivalently
so that $F[\kappa] \approx d$ in some sense. If
$d$ and $F[\kappa]$ lie in a normed subspace of $\cal{D}(\bR)$ for each
$(\bx_s,\bx_r) \in X_{sr}$, then the
norm provides a straightforward candidate for this sense of
approximation.

%In practice, recording takes place over a finite time interval
%$[0,t_d]$. For convenience, assume that this interval is the same for
%all receivers. Let $\phi \in C_0^{\infty}(0,t_d)$ be a window function
%supported in this interval. Additional constraints on $t_d$ and $\phi$
%will be be specified below.

\subsection{Hadamard's Decomposition}
The Hadamard decomposition of the acoustic Green's function provides a
basis for a norm-based formulation of FWI \cite[]{Friedlander:75,Qian:JCP24}. As presented by
\cite{Friedlander:75}, for instance, this construction applies to
second-order wave equations, but the system \ref{eqn:awe} is
equivalent fo a second-order problem: differentiate the first equation
with respect to $t$, eliminate $\bv$ via the second equation, and
rearrange to obtain
\begin{equation}
  \label{eqn:awe2}
  \frac{\rho}{\kappa}\frac{\partial^2 p}{\partial t^2} - \nabla^2 p -
  \nabla \log \rho \cdot \nabla p = \rho \frac{\partial f}{\partial t}
  \delta(\bx-\bx_s), \nonumber \\
  p=0, \, t \ll 0.
\end{equation}
If $(p,\bv)$ is a distribution solutions of the system
\ref{eqn:awe} then $p$ is a solution of equation
\ref{eqn:awe2}, as follows by differentiating the first equation of
\ref{eqn:awe} with respect to $t$ and eliminating $\bv$ via the second
equation. Conversely, if $p$ is a solution of the second-order initial
value problem\ref{eqn:awe2}, it follows from the Hadamard
decomposition detailed below and condition F1
that 
\[
  \bv = \beta \int_{-\infty}^t p
\]
has an
integrable singularity at $\bx=\bx_s$, hence 
defines a distribution, and that $(p,\bv)$ solves the system \ref{eqn:awe}.

Hadamard's construction is local, in both space-time and in the set of
acoustic models. A {\em normal neighborhood} of
$\bx_s \in \bR^3$ is an open subset of $\bR^3$ 
within which every point is connected to $\bx_s$ by  a unique ray of
geometric acoustics (a geodesic for the Riemannian metric associated
to the equation \ref{eqn:awe2}) (\cite{Friedlander:75},
p. 16). Hadamard's decomposition holds in a space-time cylinder
$\Omega \times \bR$,  and $\Omega\subset \bR^3$ 
and $M \subset (C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3))^2$ are bounded open sets satisfying 
\begin{itemize}
  \item[G1. ] for every $(\bx_s,\bx_r) \in X_{sr}$, $\bx_s, \bx_r \in \Omega$;
  \item[G2. ] for every source point $\bx_s\in X_{s}$ and $\kappa \in M$, $\Omega$ is a
    normal neighborhood of $\bx_s$.
  \item[G3. ] for every source point $\bx_s \in X_{s};$
    $\kappa \in M$, ray ${\bf X}:\bR^+\rightarrow \bR^3$
    with ${\bf X}(0)=\bx_s$ and any $t>0$, either $\{{\bf X}(s): 0\le s \le t\} \subset \Omega$
    or $\{{\bf X}(s): t < s < \infty\} \subset \bR^3 \setminus \Omega$.
  \end{itemize}
Condition G3 (the ``no return'' condition) means that once a ray
leaves $\Omega$, it does not return to $\Omega$. Conditions G1 - G3
hold if $\kappa$ is constant, and for small perturbations and some
large perturbations of constant $\kappa$: these are open conditions in
the topology of $M$.

The arguments to follow depend an additional property of solutions of
the wave equation \ref{eqn:awe2}:
\begin{itemize}
\item[G4. ] For any $n > 0$, $f \in {\cal E}'(\bR)$, there exists $K, T$ and $\delta>0$ so
  that for any 4-multiindex ${\bf m}$ of order $\le n$, $(\bx_s,\bx_r) \in
 X_{sr}, t \in \bR$,
  and any $\kappa \in M$, the solution $p$ of \ref{eqn:awe2} is smooth
  for $t>T$ and satisfies
  \[
    |D^{\bf m}p(\bx_r,t;\bx_s)| \le K e^{-\delta t}.
  \]
\end{itemize}
Condition G4 is a special case of {\em local energy decay}, a much
studied topic. Properties which imply G4 are in turn implied by the
{\em non-trapping} hypothesis on the model $\kappa \in M$: all rays of
geometric asymptotics escape to infinity (\cite{Hristova:09}, Theorem
2, \cite{EgorovShubin}, Theorem 2.104, p. 192).

Under conditions G1 - G4, the construction described by
\cite{Friedlander:75} produces for each $m \in M$ functions $\tau[\kappa]$ (travel time), $a[\kappa]$
(geometric amplitude) on $X_s \times \Omega$ and $b[\kappa]$
(remainder kernel) on $X_s \times \Omega \times \bR$, satisfying
\begin{itemize}
\item[H1. ] $\tau[\kappa], a[\kappa],$ and $b[\kappa]$ are smooth off the diagonal,
that is, $\tau[\kappa](\bx_s,\cdot) \in C^{\infty}(\Omega \setminus \{\bx_s\})$
and similarly for $a[\kappa],b[\kappa]$;
\item[H2. ] there exists $A>0$ so that for any $\kappa \in M$
  and $(\bx_s,\bx_r) \in X_{sr}$,
  $|\log a[\kappa](\bx_s,\bx_r)| < A$;
\item[H3. ] there exist $B \ge 0,\delta > 0$ so that for
$(\bx_s,\bx_r) \in X_{sr}$, and $\kappa \in
M$,
\begin{eqnarray}
  |b[\kappa](\bx_r,t;\bx_r,s)| &\le& B e^{-\delta t} \nonumber \\
  |\partial_t b[\kappa](\bx_r,t;\bx_r,s)| &\le& B e^{-\delta t}
  \label{eqn:best}
\end{eqnarray}
\end{itemize}
(The exponential decay estimate in item H3 follows from hypothesis G4.)

The Hadamard
decomposition of the Green's function $G[\kappa]$, i.e. the solution of the
equation \ref{eqn:awe2} for $f = \beta H$, is::
\begin{equation}
  \label{eqn:green}
  G[\kappa](\bx_s,\bx,t) = a[\kappa](\bx_s,\bx)\delta(t - \tau[\kappa](\bx_s,\bx)) +
  b[\kappa](\bx_s,\bx,t-\tau[\kappa])H(t - \tau[\kappa](\bx_s,\bx))
\end{equation}
%Under conditions G1 - G3 above, this decomposition is the restriction of
%the Greeen's function to $\Omega \times (-\infty,t_h]$. This fact
%implies a decomposition of $F[\kappa]$, provided that the convolution $G*f$
%restricted to $[0,t_d]$ is determined by the restriction of $G$ to
%$\Omega \times [0,t_h]$. That is the case if $t_h > t_d + |\mbox{supp
%|f/2$ (the factor of $1/2$ because $f$ is odd).
%Making this assumption (which is weaker than assumptions on $t_h$ to be made below),
%\[
%   F[\kappa](\bx_s,\bx_r)(t) =
%   \phi(t)a(\bx_s,\bx_r)\rho_s\frac{df}{dt}(t-\tau[\kappa](\bx_s,\bx_r))
% \]
%\begin{equation}
%  \label{eqn:fwdgreen}
% + \phi(t)\rho_s\int
%  \,ds \, \frac{df}{dt}(t-s) b[\kappa](\bx_s,\bx_r,s-\tau[\kappa](\bx_s,\bx_r))H(s-\tau[\kappa](\bx_s,\bx_%r))
%\end{equation}

\subsection{FWI via Least Squares}
Suppose that the source wavelet $f$ satisfies
\begin{itemize}
\item[F1. ] $f \in H^1_0(\bR)$, $\mbox{supp } \subset [-t_w,t_w]$;
\item[F2. ] $f$ is odd, $f(t) = -f(-t)$.
\end{itemize}
Then the Hadamard decomposition 
(equation \ref{eqn:green}) together with local
energy decay (assumption G4) and Young's inequality  implies that
\begin{equation}
  \label{eqn:fwdgreen}
  F[\kappa] = G[\kappa] * \rho \partial_t f \in L^p(X_{sr} \times \bR)
\end{equation}
for any $p \ge 1$, and that
the corresponding estimates are uniform in $\kappa \in M$:
there exists $C>0$ so that for $\kappa\in M$, $p \ge 1$,
\begin{equation}
  \label{eqn:fwdnorm}
  \|F[\kappa]\|_{L^p(X_{sr} \times \bR)} \le C \left\|\frac{df}{dt}\right\|_{L^p(\bR)}.
\end{equation}
%\[
%  \| \cdot \|_d = \| \cdot \|_{L^2(X_{sr} \times \bR)}
%\]

This observation justfies posing FWI as a least-squares problem: given $d \in
L^2(X_{sr}\times \bR)$, $f \in H^1_0(\bR)$, and $\rho_s>0$, find $\kappa$ to minimize
\begin{equation}
  \label{eqn:fwi}
  J_{\rm FWI}[m;d]= \frac{1}{2}\|F[\kappa]-d\|_{L^2(X_{sr} \times \bR)},
\end{equation}
%in which we have introduced the abbreviation
%\[
%  \| \cdot \|_d = \| \cdot \|_{L^2(X_{sr} \times \bR)}
%\]
Practical versions of least-squares FWI include various types of regularization and
imposition of constraints on $m$, see the references cited in the
Introduction.
%As mentioned there, local iterative optimization methods
%are the only practical approach to minimizing $J_{\rm FWI}$ due to the
%computational cost of approximating the forward map $F$ and related
%operators, but these tend to stagnate at highly sub-optimal
%(``cycle-skipped'') estimates of $m$ unless the initial estimate is of
%quite high quality, often available only at considerable cost in data
%collection and processing.

@@

Note that in the representation \ref{eqn:fwdgreen}, only the
derivative $\partial_tf$ appears. For convenience, introduce
\begin{equation}
  \label{eqn:wavelet}
  w = \rho \partial_t f
\end{equation}
The properties F1 and F2 assumed for $f$ are equivalent to these
properties of $w$:
\begin{itemize}
\item[W1. ] $w \in L^2_0(\bR)$, $\mbox{supp }w \subset [-t_w,t_w]$;
\item[W2. ] $w$ is even;
\item[W3. ] $\int \, w = 0$.
\end{itemize}
I will also
refer to $w$ as ``the wavelet'', and view $F$ as parametrized by $w$.

Introduce asymptotics in this setting by choosing $w = \wl$
from a family $\{\wl:\}$ indexed by the dimensionless parameter
$0 <\lambda \le 1\}$, and defined by
\begin{equation}
  \label{eqn:wfam}
  \wl(t) = \lambda^{-1/2}w_1(t/\lambda).
\end{equation}
The  ``mother wavelet'' $w_1$ is square-integrable of compact support
even, and mean-zero. So that property W1 holds for the entire family, 
\begin{equation}
  \label{eqn:wdiam}
  \mbox{supp }w_1 \subset [-t_w,t_w].
\end{equation}
All wavelets $\wl$ in this family
inherit these properties from $w_1$, along with its dimensions
(energy/time$^2$). It follows from these assumptions that the Fourier transform $\fwl$ is
real-valued, even, of class $H^1(\bR)$. The corresponding source waveform $f_{\lambda}$ is
given by
\begin{equation}
  \label{eqn:flambda}
  f_{\lambda}(t) = \int_{-\infty}^t\wl =
  \lambda^{1/2}f_1\left(\frac{t}{\lambda}\right).
\end{equation}
Note that for $p \ge 1$,
\begin{equation}
  \label{eqn:lambdanorms}
  \|\wl\|_{L^p(\bR)} = \lambda ^{\frac{1}{p}-\frac{1}{2}}\|w_1\|_{L^p(\bR)},\,\|f_{\lambda}\|_{L^p(\bR)}
  = \lambda ^{\frac{1}{p}+\frac{1}{2} \|f_1\|_{L^p(\bR)}}.
\end{equation}

@@

The modeling operator with wavelet $\wl$ is denoted $\Fl$. From the
decomposition \ref{eqn:fwdgreen}, $\Fl$ is the sum of two terms:
\begin{equation}
  \label{eqn:Flamdecomp}
  \Fl = \Fla + \Flb.
\end{equation}
where
\begin{equation}
  \label{eqn:Fl0}
  \Fla[\kappa](\bx_r,t;\bx_s) =
  a[\kappa](\bx_s,\bx_r)\wl(t-\tau[\kappa](\bx_s,\bx_r))
\end{equation}
and
\[
\Flb[\kappa] = \rho_s\int
   \,ds \, \frac{df_{\lambda}}{dt}(t-s) b[\kappa](\bx_s,\bx_r,s-\tau[\kappa](\bx_s,\bx_r))H(s-\tau[\kappa](\bx_s,\bx_r))
\]
\[
  =\rho_s
  f_{\lambda}(t-\tau(\bx_s,\bx_r))b[\kappa](\bx_s,\bx_r,0)
\]
\begin{equation}
  \label{eqn:Fl1}
    -
    \rho_s\int
   \,ds \, f_{\lambda}(t-s) (\partial_t
   b[\kappa](\bx_s,\bx_r,s-\tau[\kappa](\bx_s,\bx_r)))H(s-\tau[\kappa](\bx_s,\bx_r))
\end{equation}
From hypothesis G4, the bound \ref{eqn:lambdanorms} , and Young's inequality, the $\|\Fl1[\kappa]\|_{ L^p(X_{sr} \times \bR)}$ is bounded by a multiple of
$\lambda^{\frac{1}{p}+\frac{1}{2}}$, uniformly over $\kappa \in M$.
I will write ``$+ \Ol$'' to mean an  error uniformly bounded in $L^2$ by
a multiple of $\lambda$ over $m \in M$. Thus
\begin{equation}
  \label{eqn:fwdapprox}
  \Fl[\kappa](\bx_s,\bx_r) = \Fla[\kappa] +   \Ol
\end{equation}

\subsection{AWI via Optimization}
